---
title: "EDA_Suppl&Financial_Accounts"
author: "Avinash"
date: "2023-10-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data into a variable and check the strucutre of the data.
```{r}

# Load the data set into a variable
bops_data <- read.csv("BalanceOfPayment.csv")

# Structure of the data
str(bops_data)
```

## Converting the requisite columns into factor variables.
```{r}

# Convert the requisite columns into factor 
cols_to_convert <- c("Country", "Code", "Indicator", "Icode")

# Use lapply() function to apply the factor() function to each specified column in `bops_data` data frame.
bops_data[cols_to_convert] <- lapply(bops_data[cols_to_convert], factor)

# Check the structure of the data again to confirm the changes
str(bops_data)
```

## Filter out the data set to have the following necessary indicators: Supplementary Accounts, and Financial Accounts.
```{r}

# load the `dplyr` package into R session
library(dplyr)

# List of indicators to filter
indicators_to_filter <- c("Supplementary Items, Reserve Assets (with Fund Record), US Dollars", 
"Supplementary Items, Total Current + Capital Account, US Dollars",
"Supplementary Items, Reserve Position in the Fund (with Fund Record), US Dollars",
"Supplementary Items, Portfolio Investment, Net Incurrence of Liabilities (Excluding Exceptional Financing), US Dollars", 
"Financial Account, Portfolio Investment, Net Acquisition of Financial Assets, US Dollars",
"Financial Account, Portfolio Investment, Net Acquisition of Financial Assets, Equity and Investment Fund Shares, US Dollars",
"Net Lending (+) / Net Borrowing (-) (Balance from Current and Capital Account), US Dollars",
"Financial Account, Other Investment, Loans, US Dollars",
"Financial Account, Portfolio Investment, Net Acquisition of Financial Assets, Debt Securities, US Dollars")

# Filter the data
relevant_financial_metrics <- bops_data %>% 
  filter(Indicator %in% indicators_to_filter)

# Structure of the data
str(relevant_financial_metrics)

# Print the first and last 10 rows of the data
head(relevant_financial_metrics, 15)
tail(relevant_financial_metrics, 10)
```

## SMART Question -1:

How have the global trends evolved for the "Total Current + Capital Account" from 2015 to 2020. (Supplementary Account)

```{r}

# Load the necessary libraries
library(ggplot2)
library(tidyr)

# Filter the dataset for the 'Total Current + Capital Account' indicator
total_current_capital_account <- relevant_financial_metrics[relevant_financial_metrics$Indicator == "Supplementary Items, Total Current + Capital Account, US Dollars", ]

# Reshape the data to long format
long_data <- total_current_capital_account %>% 
  pivot_longer(cols = starts_with("X"), names_to = "Year", values_to = "Value")

# Convert the "XYYYY" format to just "YYYY"
long_data$Year <- as.numeric(sub("X", "", long_data$Year))

# Time series plot
ggplot(long_data, aes(x=Year, y=Value, group=Country)) + 
  geom_line(aes(color=Country), alpha=0.5) + 
  labs(title="Evolution of Total Current + Capital Account from 2015 to 2020", y="Total Current + Capital Account", x="Year") + 
  theme(legend.position="none") # Hide legend for clarity

```

### By plotting time series data for many countries on a single graph, the visualization became cluttered, making it difficult to discern patterns. Hence, a more effective strategy would be to segregate the countries into two categories: "Top 10 countries by economy" and "Top 10 emerging countries."

```{r}

# top 10 countires by economy
top_10_Economic_Countries <- c("United States", "China, P.R.: Mainland", "Japan", "Germany", "India", "United Kingdom", 
                               "France", "Russian Federation", "Canada", "Italy")

# top 10 emerging countries
top_10_emerging_countries <- c("Brazil", "Russian Federation", "India", "China, P.R.: Mainland", "South Africa",
                               "Mexico", "Indonesia", "Türkiye, Rep. of", "Nigeria", "Philippines")

# Load the necessary libraries
library(ggplot2)
library(tidyr)
library(RColorBrewer)

# Filter data for the top 10 economies
top_10_economies_data <- subset(long_data, Country %in% top_10_Economic_Countries)

# Plot for top 10 economies
plot_top_10_economies <- ggplot(top_10_economies_data, aes(x=Year, y=Value, group=Country, color=Country)) + 
  geom_line(linewidth=1) + 
  labs(title="Evolution of Total Current + Capital Account from 2015 to 2020\nfor Top 10 Economies", 
       y="Total Current + Capital Account", 
       x="Year") + 
  theme(legend.position="right", 
        plot.title=element_text(hjust=0.5))

# Filter data for the top 10 emerging economies
top_10_emerging_data <- subset(long_data, Country %in% top_10_emerging_countries)

# Plot for top 10 emerging economies
plot_top_10_emerging <- ggplot(top_10_emerging_data, aes(x=Year, y=Value, group=Country, color=Country)) + 
  geom_line(linewidth=1) + 
  labs(title="Evolution of Total Current + Capital Account from 2015 to 2020\nfor Top 10 Emerging Economies", 
       y="Total Current + Capital Account", 
       x="Year") + 
  scale_color_brewer(palette="Set3", name="Emerging Economies") + 
    theme(legend.position="right", 
        plot.title=element_text(hjust=0.5))

# Display the plots
plot_top_10_economies
plot_top_10_emerging
```

## SMART question - 2
How does the distribution of 'Net Acquisition of Financial Assets' look like for countries in 2020? Are there outliers?

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Extract the data for the 'Net Acquisition of Financial Assets' in 2020
data_2020 <- relevant_financial_metrics$X2020[relevant_financial_metrics$Indicator == "Financial Account, Portfolio Investment, Net Acquisition of Financial Assets, US Dollars"]

# Apply a log transformation to the data; adding a small constant to handle zeros or negative values
log_data_2020 <- log(data_2020 + max(abs(data_2020))*1.01) 

# Calculate statistics for the transformed data
mean_log_value <- mean(log_data_2020, na.rm = TRUE)
median_log_value <- median(log_data_2020, na.rm = TRUE)

# Plotting the histogram
ggplot(data.frame(Value=log_data_2020), aes(x=Value)) +
  geom_histogram(binwidth=0.05, fill="blue", alpha=0.7) +  
  geom_vline(aes(xintercept=mean_log_value), color="red", linetype="dashed", size=1) +
  geom_vline(aes(xintercept=median_log_value), color="green", linetype="dashed", size=1) +
  annotate("text", x = mean_log_value + 1, y = 30, label = paste("Mean:", round(mean_log_value, 2))) +
  annotate("text", x = median_log_value + 1, y = 20, label = paste("Median:", round(median_log_value, 2))) +
  labs(title="Log-transformed Distribution of Net Acquisition of Financial Assets in 2020",
       x="Log-transformed Net Acquisition of Financial Assets",
       y="Number of Countries") + 
  theme(legend.position="right", 
        plot.title=element_text(hjust=0.5))

# Plotting the boxplot
ggplot(data.frame(Value=log_data_2020), aes(x="", y=Value)) +
  geom_boxplot(fill="lightblue", color="blue", outlier.color="red", outlier.shape=16) +
  labs(title="Boxplot of Log-transformed Net Acquisition of Financial Assets in 2020",
       y="Log-transformed Net Acquisition of Financial Assets") + 
  theme(legend.position="right", 
        plot.title=element_text(hjust=0.5))


# Determine the 1st and 3rd quartiles and IQR
Q1 <- quantile(log_data_2020, 0.25)
Q3 <- quantile(log_data_2020, 0.75)
IQR_value <- IQR(log_data_2020)

# Using ggplot2 to create a boxplot with customized y-axis limits
ggplot(data.frame(Value=log_data_2020), aes(x="", y=Value)) +
  geom_boxplot(fill="lightblue", color="blue", outlier.shape=NA) +
  geom_hline(yintercept = median(log_data_2020), linetype="dashed", color = "red", size=0.5) +  # Median line
  labs(title="Boxplot of Log-transformed Net Acquisition of Financial Assets in 2020",
       y="Log-transformed Net Acquisition of Financial Assets") + 
  theme_minimal() +
  coord_cartesian(ylim=c(Q1 - 0.5*IQR_value, Q3 + 0.5*IQR_value))  
```


## SMART question - 3
Do countries with higher average 'Equity and Investment Fund Shares' over the period 2015-2020 also possess higher average 'Debt Securities' during the same period?

```{r}
# filter the data for the two indicators
equity_data <- relevant_financial_metrics[relevant_financial_metrics$Indicator == "Financial Account, Portfolio Investment, Net Acquisition of Financial Assets, Equity and Investment Fund Shares, US Dollars", ]
debt_data <- relevant_financial_metrics[relevant_financial_metrics$Indicator == "Financial Account, Portfolio Investment, Net Acquisition of Financial Assets, Debt Securities, US Dollars", ]

# Calculate the mean for the years 2015-2020 for each country
equity_data$mean_equity <- rowMeans(equity_data[,c("X2015", "X2016", "X2017", "X2018", "X2019", "X2020")], na.rm = TRUE)
debt_data$mean_debt <- rowMeans(debt_data[,c("X2015", "X2016", "X2017", "X2018", "X2019", "X2020")], na.rm = TRUE)

# Merge the datasets by country
merged_data <- merge(equity_data[,c("Country", "mean_equity")], debt_data[,c("Country", "mean_debt")], by="Country")

# Calculate the correlation
correlation <- cor(merged_data$mean_equity, merged_data$mean_debt, use="complete.obs")

correlation
```


```{r}
# Load necessary libraries
library(ggplot2)

# filter the data for the two indicators
equity_data <- relevant_financial_metrics[relevant_financial_metrics$Indicator == "Financial Account, Portfolio Investment, Net Acquisition of Financial Assets, Equity and Investment Fund Shares, US Dollars", ]
debt_data <- relevant_financial_metrics[relevant_financial_metrics$Indicator == "Financial Account, Portfolio Investment, Net Acquisition of Financial Assets, Debt Securities, US Dollars", ]

# Calculate the mean for the years 2015-2020 for each country
equity_data$mean_equity <- rowMeans(equity_data[,c("X2015", "X2016", "X2017", "X2018", "X2019", "X2020")], na.rm = TRUE)
debt_data$mean_debt <- rowMeans(debt_data[,c("X2015", "X2016", "X2017", "X2018", "X2019", "X2020")], na.rm = TRUE)

# Merge the datasets by country
merged_data <- merge(equity_data[,c("Country", "mean_equity")], debt_data[,c("Country", "mean_debt")], by="Country")

# Create a new column to identify whether the country is a top economy or emerging
merged_data$Type <- ifelse(merged_data$Country %in% top_10_Economic_Countries, "Top Economy", "Emerging Country")

# Scatter plot differentiating between the top 10 economies and the top 10 emerging countries
ggplot(merged_data, aes(x=mean_equity, y=mean_debt, color=Type)) + 
  geom_point(aes(shape=Type), size=3) +  
  geom_smooth(method="lm", se=FALSE, color="red", linetype="dashed") +  # Add a linear regression line
  labs(title="Average Equity vs. Average Debt Securities (2015-2020)",
       x="Average Equity and Investment Fund Shares",
       y="Average Debt Securities") +
  theme(legend.position="right", plot.title=element_text(hjust=0.5)) +
  scale_color_manual(values=c("Top Economy"="blue", "Emerging Country"="red")) +
  scale_shape_manual(values=c("Top Economy"=19, "Emerging Country"=17))  

```

## SMART question - 4
What is the variance in the average 'Reserve Position in the Fund' across all countries over the years 2015-2020?

```{r}
# Load necessary libraries
library(ggplot2)

# Filter data for 'Reserve Position in the Fund'
reserve_data <- relevant_financial_metrics[relevant_financial_metrics$Indicator == "Supplementary Items, Reserve Position in the Fund (with Fund Record), US Dollars", c("X2015", "X2016", "X2017", "X2018", "X2019", "X2020")]

# Calculate the average for each country over the years 2015-2020
reserve_data$mean_reserve <- rowMeans(reserve_data, na.rm = TRUE)

# Calculate variance
variance <- var(reserve_data$mean_reserve, na.rm = TRUE)

# Plot Histogram
hist(reserve_data$mean_reserve, main="Distribution of Average Reserve Position in the Fund (2015-2020)", xlab="Average Reserve Position in the Fund", breaks=50)

# Plot Boxplot
boxplot(reserve_data$mean_reserve, main="Boxplot of Average Reserve Position in the Fund (2015-2020)", ylab="Average Reserve Position in the Fund")

```

## SMART question - 5
Hypothesis Testing
Hypothesis: The average 'Total Current + Capital Account' is equal to the average 'Net Acquisition of Financial Assets' for countries in 2020.

```{r}

# Extracting data
supplementary_accnts_data <- relevant_financial_metrics[relevant_financial_metrics$Indicator == "Supplementary Items, Total Current + Capital Account, US Dollars", ]
financial_accnts_data <- relevant_financial_metrics[relevant_financial_metrics$Indicator == "Financial Account, Portfolio Investment, Net Acquisition of Financial Assets, US Dollars", ]

# Calculating averages for 2015-2020
supplementary_accnts_data$average <- rowMeans(supplementary_accnts_data[, c("X2015", "X2016", "X2017", "X2018", "X2019", "X2020")], na.rm = TRUE)
financial_accnts_data$average <- rowMeans(financial_accnts_data[, c("X2015", "X2016", "X2017", "X2018", "X2019", "X2020")], na.rm = TRUE)

# T-test
t_test_result <- t.test(supplementary_accnts_data$average, financial_accnts_data$average)

# Print the t-test result
print(t_test_result)

# Boxplot
boxplot(supplementary_accnts_data$average, financial_accnts_data$average, names=c("Supplementary Items", "Financial Account"), main="Comparison of Averages", ylab="Value", xlab="Dataset")

```
The p-value is slightly above 0.05, which means, at the 5% significance level, we fail to reject the null hypothesis. This suggests that there isn't enough statistical evidence to say that the average 'Total Current + Capital Account' is different from the average 'Net Acquisition of Financial Assets' over the period 2015-2020.

